---
title: "Assignment 4"
author: "Natalya Shelchkova"
date: "February 23, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Section 10.5 Problem 5

What does tibble::enframe() do? When might you use it?

Enframe() takes atomic vectors/lists and makes them into two column data frames. This is useful when you load in data as a list and you want to convert it into a data frame, or you have to set variables and then you want to save the variables you set as a data frame. 

## Section 12.6.1 Problem 3

I claimed that iso2 and iso3 were redundant with country. Confirm this claim.

If we take the number of distinct values in iso2 and iso3 we will see that this number is the same:

```{r}
who <- tidyr::who
iso2_count <- n_distinct(who["iso2"])
iso3_count <- n_distinct(who["iso3"])
iso2_count
iso3_count
```

At the very least this means that they have the same number of distinct values. Furthermore, if we look at the unique values in iso2 and iso3, we will see that they are the same, with the exception that iso3 is a 3 letter abbreviation:

```{r}
iso2_unique <- unique(who["iso2"])
iso3_unique <- unique(who["iso3"])
iso2_unique
iso3_unique

```


## Section 12.6.1 Problem 4

For each country, year, and sex, compute the total number of cases of TB. Make an informative visualization of the data.

```{r, echo = FALSE}
library(tidyverse)
who <-  tidyr::who

# # Removes NA, and makes the columns new_sp_m014:newrel_f65 into variables
# who1 <- who %>% 
#   gather(new_sp_m014:newrel_f65, key = "key", value = "cases", na.rm = TRUE)
# 
# # makes names consistent
# who2 <- who1 %>% 
#   mutate(key = stringr::str_replace(key, "newrel", "new_rel"))
# 
# # seperates cases into whether the tb is new or recurrent, the type of tb, and the sax/age of the individual
# who3 <- who2 %>% 
#   separate(key, c("new", "type", "sexage"), sep = "_")
# 
# # removes the columns new, iso2, and iso3 since they are reduntant
# who4 <- who3 %>% 
#   select(-new, -iso2, -iso3)
# 
# # seperates sex and age into two columns
# who5 <- who4 %>% 
#   separate(sexage, c("sex", "age"), sep = 1)

# the pipe version of the above code
tb_data <- who %>%
  gather(code, value, new_sp_m014:newrel_f65, na.rm = TRUE) %>% 
  mutate(code = stringr::str_replace(code, "newrel", "new_rel")) %>%
  separate(code, c("new", "var", "sexage")) %>% 
  select(-new, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1)

## Computing totals

# by year
year_totals <-  tb_data %>%
  group_by(year) %>%
  summarise(year_total = sum(value))

q <- ggplot(data = year_totals) + 
  geom_point(mapping = aes(x = year, y = year_total)) +
  geom_smooth(mapping = aes(x = year, y = year_total))
q + labs(x = "Year", y = "Total Cases of TB", title = "Cases of TB Over the Years")

# by country
country_totals <- tb_data %>%
  group_by(iso2) %>%
  summarise(country_total = sum(value))

p <- ggplot(data = country_totals) + 
  geom_col(mapping = aes(x = iso2, y = country_total))
p1 <- p + labs(x = "Country", y = "Total Cases of TB", title = "Cases of TB per Country")
p1 + theme(axis.text.x = element_text(angle = 90, size = 5))

# by sex
sex_totals <- tb_data %>%
  group_by(sex) %>%
  summarise(sex_total = sum(value))

r <- ggplot(data = sex_totals) + 
  geom_col(mapping = aes(x = sex, y = sex_total))
r + labs(x = "Sex", y = "Total Cases of TB", title = "Cases of TB Based on Sex")

```

### Case Study

```{r Loading Data, echo = FALSE}

rawdat <- readRDS("~/Data Science in R - MA415/MA415-Homework4/rawdat.RDS")

bb <- read_csv("billboard.csv")
bb.1 <- bb %>% select(year, artist = artist.inverted, track, time, date = date.entered, id = 8:82)
names(bb.1)[6:80] = paste("wk",1:75, sep = "")
bb_data <- bb.1

```

## Reproducing Table 4 to Table 6

```{r}

# transformed_table <- melt(rawdat, id = c("religion"), variable.name = "income", value.name = "freq")
# OR

transformed_table_v2 <- rawdat %>% gather(key = "income", value = "freq", -religion)
transformed_table_v2

```


## Reproducing Table 7 to Table 8

```{r}

bb_v2 <- bb_data %>% gather(key = "week", value = "rank", -year, -artist, -time, -track, -date)
bb_v3 <- bb_v2 %>% separate(week, c("rem", "week"), sep = 2)
bb_v4 <- bb_v3 %>% select(year, artist, track, time, date, week, rank)
bb_v4

```
